---
params:
  researcher_ids: ""
  examination_years : ""
  user_mailaddress : ""
  researcher_name : ""
output:
  pdf_document:
    latex_engine: lualatex
    template: snsf.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preparations, include=FALSE}

# Extract parameters into variable for convenience
researcher_ids <- params$researcher_ids
examination_years <- params$examination_years
user_mailaddress <- params$user_mailaddress
researcher_name <- params$researcher_name

include_publications <- TRUE
 
if (researcher_ids == "")
  stop("No researcher ID(s) provided!")

# Package preparations
library(tidyverse)
library(here)
library(roadoi)
library(readxl)
library(magrittr)
library(httr)
library(jsonlite)
library(kableExtra)
library(lubridate)
library(scales)
library(lintr)
library(conflicted)
library(showtext)

# Package conflict preferences
conflict_prefer("here", "here")
conflict_prefer("filter", "dplyr")
conflict_prefer("extract", "magrittr")

# Load dependent functions
source(here("core", "dimensions_functions.R"))
source(here("core", "oa_detection.R"))
source(here("core", "report_generation_functions.R"))

# Set locale to display today's month in English (supports all OS)
Sys.setlocale("LC_TIME", "C")

# Mail address only in lowercase for API queries
user_mailaddress <- str_to_lower(user_mailaddress)
```


```{r load_articles, include=FALSE}
# Get articles since 2015 of the chosen researcher from Dimensions
publications <- get_researcher_publications(researcher_ids, examination_years)

# Subset articles
articles <- publications %>%  
  filter(type == "article")

# Number of NA DOIs
nrow(articles %>% filter(is.na(doi)))

# Set the variable for the title of the document (set in the end)
document_title <- paste0("Open Access Check: ", researcher_name)

# Show message if no articles were found for researcher
exit_message <- ""
if (nrow(articles) == 0) {
  exit_message <- 
    paste0("No scientific articles were found for researcher ", researcher_name,
           " for the examined years ", get_year_list_string(examination_years), 
           " in Dimensions. Since there are still issues with", 
           " the reliability of automated analyses of publication formats", 
           " other than articles, these formats could not be included.")
}

```

`r if (nrow(articles) == 0) { "
---
title: No articles found
---
" }`

`r if (nrow(articles) == 0) { exit_message }`

`r if (nrow(articles) == 0) { knitr::knit_exit() }`

```{r determine_oa_status, include=FALSE}

# Call Unpaywall with article DOIs and classify the articles 
# into the SNSF OA categories
articles <- determine_oa_status(articles, user_mailaddress)

# Remove duplicates
articles <- articles %>% 
  distinct(id, .keep_all = TRUE)

# Classify the closed articles in classes depending on time passed since 
# their publication
articles <- classify_closed_articles(articles)

# Get number of articles successfully queried to Unpaywall 
number_successfull_unpaywall_calls <- articles %>% 
  filter(unpaywall_status == "OK") %>% 
  nrow()

# Show message if none of the found articles could be queried to Unpaywall
exit_message_unpaywall <- ""
if (number_successfull_unpaywall_calls == 0) {
  exit_message_unpaywall <- 
    paste0("None of the scientific articles found for ", researcher_name,
           " in the examined years ", get_year_list_string(examination_years), 
           " could be queried successfully to Unpaywall. Concerned ", 
           " articles: ", paste(articles$doi, collapse = ", "), ".")
}

```


`r if (number_successfull_unpaywall_calls == 0) { "
---
title: Article(s) could not be queried to Unpaywall
---
" }`

`r if (number_successfull_unpaywall_calls == 0) { exit_message_unpaywall }`

`r if (number_successfull_unpaywall_calls == 0) { knitr::knit_exit() }`

```{r create_plot, include=FALSE}
# Sum up the OA classes
oa_totals <- calculate_oa_totals(articles)

# Calculate number of articles
total_n <- sum(oa_totals$n, na.rm = TRUE)

# Create plot
p_oa_colors <- generate_oa_plot(oa_totals, researcher_name)

# Create caption for code chunk
caption <-
  paste0("SNSF OA categories, ", prettyNum(total_n, big.mark = "'"), 
    " articles by ", researcher_name, ", number of articles in brackets")

# Percentage of OA (SNSF definition and broad definition) publications
articles <- articles %>% 
  # Only SNSF statuses gold green hybrid are seen as SNSF OA
  mutate(is_oa_snsf = 
           if_else(oa_status %in% c("gold", "green", "hybrid"), TRUE, FALSE))

# Create some dynamic parts of the report text
short_summary <- generate_short_summary(researcher_name, articles)
```

__\textcolor{red}{The SNSF Open Access Check is a prototype and reports like this one can still contain errors.}__

`r short_summary`

## Report of `r format(lubridate::now(), "%B %d, %Y")`

For the year(s) `r get_year_list_string(examination_years)` we found **`r nrow(publications)`** publications in Dimensions[^1] for researcher `r researcher_name`, listing her/him as author, regardless of position in the author line.

[^1]: https://app.dimensions.ai


```{r publication_types, echo=FALSE}
# Draw kable with publication types
kable(publications %>%
        count(type) %>%
        arrange(type) %>%
        mutate(type = paste(str_to_upper(str_sub(type, 1, 1)),
                            str_sub(type, 2, nchar(type)),
                            sep = "")) %>%
        mutate(`Used in OA Check` = if_else(type == "Article", "Yes", 
               "No")) %>% 
        rename(`Publication Type (Dimensions)` = type), 
      format = "latex") %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(position = "center")
```

__Only the `r nrow(publications %>% filter(type == "article"))` articles found in Dimensions are analysed.__ Lack of coverage of other publication formats does not allow for a useful analysis beyond articles.

`r number_successfull_unpaywall_calls` (`r round((number_successfull_unpaywall_calls / nrow(articles)) * 100, 1)`%) of these articles have DOIs available and could be queried successfully to Unpaywall[^2] in order to obtain OA metadata. Articles are then classified into OA categories according to SNSF's open access definitions. For details please refer to the last sections of this report. 

[^2]: https://unpaywall.org

```{r plot_oa, echo=FALSE, fig.cap=caption, fig.width=6.5, fig.height=3.6, fig.align='center', fig.showtext=TRUE}

# Add SNSF Font
font_add(family = "Source Sans Pro", 
         regular = here("report", "font", "SourceSansPro-Regular.ttf"))

# Activate showtext
showtext_auto()

p_oa_colors
```

`r if(include_publications){"\\newpage"}`

`r if(include_publications){"# Appendix"}`

`r if(include_publications & nrow(filter(articles, oa_status == "closed")) > 0){"## Closed articles  "}`

```{r appendix_closed, echo=FALSE, results='asis'}
# Print the table with the articles used in the calculation for the OA shares
if (include_publications) {
   
  # Bring data to correct format for displaying it in table
  table_data <- articles %>%
    # When Dimensions does not have the journal name of this article, fall 
    # back to Unpaywall journal title (from Crossref) when available
    mutate(journal_name = ifelse(is.na(journal_name) & !is.na(journal$title), 
                                 journal$title, journal_name)) %>% 
    select(title.x, doi, closed_oa_class, journal_name, date, oa_status,
           months_since_pub, version, journal_is_oa, host_type, is_oa, 
           pdf_link) %>%
    mutate(date = if_else(oa_status == "closed",
                          # Dimensions dates could also be not full dates
                          if_else(nchar(date) == 10,
                                  # Only calculate days since publication,
                                  # when it is a full date
                                 paste0(date, " (", closed_oa_class, 
                                         " months)"),
                                  date), # No exact pub date from Dimensions
                          date),
           title.x = paste0(title.x, " (", doi, ")")) %>% 
    # Change order: First the closed, then the normal order
    mutate(oa_status = factor(
      articles$oa_status,
      c("closed", "gold", "green", "hybrid", "other OA")
    )) %>%
    # Order by the age of the article
    arrange(months_since_pub) %>%
    # Remove some variables not needed to be displayed separately
    select(-doi, -months_since_pub) %>%
    rename(Title = title.x,
           Journal = journal_name,
           `Publication date` = date,
           `SNSF OA Status` = oa_status) %>%  
    # Escape special Latex characters, so the argument escape = FALSE can be 
    # set for kable(), which is needed to allow links
    mutate(Title = map_chr(Title, function(x) {
      sanitize_tex_string(x)
    }),
    Journal = map_chr(Journal, function(x) {
      sanitize_tex_string(x)
    })) 
  
  # If there are closed articles, show them in first table
  if (nrow(filter(articles, oa_status == "closed")) > 0) {
    # Subset the closed articles
    table_subset <- table_data %>%
                   filter(`SNSF OA Status` == "closed")
    
    for (idx in 1:length(levels(table_subset$closed_oa_class))) {
      # Get the closed class
      closed_class <- levels(table_subset$closed_oa_class)[idx]
      
      # When it actually is a closed class with data subset and show articles
      if (!(closed_class %in% c("open", "undefined")) &
          nrow(filter(table_subset, closed_oa_class == closed_class)) > 0) {
        
        # Draw kable with subsetted articles
        kab <- kable(table_subset %>%
                       # Subset this closed class
                       filter(closed_oa_class == closed_class) %>%  
                       select(-is_oa, -closed_oa_class, -version, 
                              -journal_is_oa, -host_type, -pdf_link), 
                     format = "latex", longtable = TRUE) %>%

          column_spec(1, "7cm") %>%
          column_spec(2, "3.5cm") %>%
          column_spec(3, "2cm") %>%
          column_spec(4, "1.5cm") %>%
          row_spec(0, bold = TRUE) %>%
          kable_styling(latex_options = c("hold_position", "repeat_header"))
        
        # Cosmetic changes
        closed_class_print <-
          case_when(closed_class == "<= 6" ~ "$\\le$ 6",
                    closed_class == "> 24" ~ "$>$ 24",
                    TRUE ~ closed_class)
        
        # Print closed class as title and show kable (and do not break)
        cat(paste0("### Time since publication: ", closed_class_print,
                   " months "))
        cat(kab)
      }
    }
  }
}
```

`r if(include_publications & nrow(filter(articles, oa_status == "other OA")) > 0){"## Other OA articles  "}`

`r if(include_publications & nrow(filter(articles, oa_status == "other OA")) > 0){"*SNSF requires articles to be at least available as published versions or author's accepted manuscripts (including revisions after peer review, excluding publisher's layout). If the data obtained from Unpaywall does not allow to determine if an article fulfills this criterion, the article is considered \"other OA\".*\\newline"}`

```{r appendix_other_oa, echo=FALSE}
# Print the table with the articles used in the calculation for the OA shares
if (include_publications) {
  # If there are other OA articles, show them here
  if (nrow(filter(articles, oa_status == "other OA")) > 0) {
    # Subset the other OA articles
    table_subset <- table_data %>%
      filter(`SNSF OA Status` == "other OA") %>%
      # Create new variable with the reason for the article being other OA 
      mutate(reason = if_else(is_oa == TRUE & # Clarity purposes, anyway T
                                host_type == "repository" & 
                                version == "NULL", 
                              "In repository, but no version description found", 
                              "")) %>% 
      mutate(reason = 
               if_else(reason == "", 
                       if_else(is_oa == TRUE & # For clarity, always TRUE
                                 host_type == "repository" & 
                                 !(version %in% c("publishedVersion",
                                                  "acceptedVersion", 
                                                  "updatedVersion"
                                 )), 
                               paste0("not peer-reviewed, '", version,
                                      "'"), 
                               ""), 
                       reason)) %>%
      mutate(reason = if_else(reason == "", "Unknown reason", reason)) %>%
      # Add the link to the found PDF location
      mutate(`SNSF OA Status` = paste0("\\href{", pdf_link, "}{\\underline{",
                                       `SNSF OA Status`,
                                       "}}")) %>%
      mutate(`SNSF OA Status` = paste0(`SNSF OA Status`, " (", reason, ")")) %>%
      select(-is_oa, -closed_oa_class, -version, -journal_is_oa, -host_type, 
             -reason, -pdf_link)
    # Draw kable with subsetted articles
    kable(table_subset, format = "latex", escape = FALSE, longtable = TRUE) %>%
      # Adjust longtable options
      kable_styling(latex_options = c("hold_position", "repeat_header")) %>%
      column_spec(1, "6cm") %>%
      column_spec(2, "3cm") %>%
      column_spec(3, "2cm") %>%
      column_spec(4, "3cm") %>%
      row_spec(0, bold = TRUE)
  }
}

```

`r if(include_publications & nrow(filter(articles, is_oa_snsf == T)) > 0){"## Open articles (according to SNSF definitions)  "}`

```{r appendix_open, echo=FALSE}
# Print the table with the articles used in the calculation for the OA shares
if (include_publications) {
  # If there are closed articles, show them in first table
  if (nrow(filter(articles, is_oa_snsf == TRUE)) > 0) {
    # Subset the closed articles
    table_subset <- table_data %>%
      filter(`SNSF OA Status` %in% c("gold", "green", "hybrid")) %>%
      # Add the link to the found PDF location
      mutate(`SNSF OA Status` = paste0("\\href{", pdf_link, "}{\\underline{", 
                                       `SNSF OA Status`, 
                                       "}}")) %>%  
      select(-is_oa, -closed_oa_class, -version, -journal_is_oa, 
             -host_type, -pdf_link)
      
    # Draw kable with subsetted articles
    kable(table_subset, format = "latex", escape = FALSE, longtable = TRUE) %>%
      # Adjust longtable options
      kable_styling(latex_options = c("hold_position", "repeat_header")) %>%
      column_spec(1, "7cm") %>%
      column_spec(2, "3.5cm") %>%
      column_spec(3, "2cm") %>%
      column_spec(4, "1.5cm") %>%
      row_spec(0, bold = TRUE)
  }
}
```

\newpage

```{r methods, child = here("text", "methods.md")}
```

```{r limitations, child = here("text", "limitations.md")}
```

```{r howto, child = here("text", "howto.md")}
```

---
title: "`r document_title`"
---